### 2013-06-07 papers/sle2013/sle2013.tex
sle code examples

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage[utf8]{inputenc}
\usepackage[unicode=true,hidelinks]{hyperref}

\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\include{macros}

\newcommand{\ie}{\emph{i.e.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\cf}{\emph{cf.~}}
\newcommand{\etal}{\emph{et al.~}}
\newcommand{\etc}{\emph{etc.}}
\newcommand{\aka}{\emph{a.k.a.}}

\begin{document}

\mainmatter

\title{Using Dependent Types to Build Safe and Concise JavaScript Foreign Function Interfaces}
\titlerunning{Dependent Types Rock}

\author{Julien Richard-Foy \and Olivier Barais\and Jean-Marc Jézéquel}
\authorrunning{Julien Richard-Foy \emph{et. al.}}

\institute{IRISA, Université de Rennes 1, France. \texttt{\{first\}.\{last\}@irisa.fr}}

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
Several programming languages can target JavaScript as a back-end, giving developers programming language features
that are absent from JavaScript, such as static typing. However, the Web browser APIs, which are needed to interact
with a Web page, are designed for JavaScript, making it challenging to expose them in a statically typed language.
Indeed, existing statically typed languages exposing Web browser APIs either break type safety or give developers
less control than if they were using JavaScript. This article shows how to expose Web browser APIs in Scala in a
type safe way while keeping the same level of control as with native APIs by using dependent types and functional
dependencies. We validate this approach in designing safe and concise foreign function interfaces between Scala and
JavaScript for DOM events handling and DOM manipulation. We compared this approach to common frameworks such as GWT,
Fay, Kotlin, Dart and SharpKit.
\keywords{Dependent Types, JavaScript, Scala, Foreign Function Interface}
\end{abstract}

\section{Introduction}

Web applications are attractive because they require no installation or deployment step on clients and enable large
scale collaborative experiences. However, writing large Web applications is known to be
difficult~\cite{Mikkonen08_SpaghettiJs,Preciado05_RIAMethodologyNecessity}. One challenge comes from the fact that
the JavaScript programming language -- which is currently the only action language natively supported by almost all
Web clients -- lacks of constructs making large code bases maintainable (\eg static typing, first-class modules).

One solution consists in considering JavaScript as an assembly language\footnote{\cf
\href{http://asmjs.org/}{http://asmjs.org/}} and generating JavaScript from compilers of full-featured and
cutting-edge programming languages. Incidentally, an increasing number of programming languages or compiler backends
can generate JavaScript code (\eg Java/GWT~\cite{Chaganti07_GWT},
SharpKit\footnote{\href{http://sharpkit.net}{http://sharpkit.net}}, Dart~\cite{Griffith11_Dart},
Kotlin\footnote{\href{http://kotlin.jetbrains.org/}{http://kotlin.jetbrains.org/}},
ClojureScript~\cite{McGranaghan11_ClojureScript}, Fay\footnote{\href{http://fay-lang.org/}{http://fay-lang.org/}},
Haxe~\cite{Cannasse08_HaXe}, Opa\footnote{\href{http://opalang.org/}{http://opalang.org/}}).

However, compiling to JavaScript is not enough. Developers also need the Web browser programming environment: they
need to interact with the Web page, to build DOM fragments, to listen to user events, \etc. A Foreign Function
Interface mechanism could be used to make browser’s APIs available to the developers. However, JavaScript APIs are
not statically typed and make a heavy use of overloading, making them hard to expose in a statically typed language.

Indeed, existing statically typed languages compiling to JavaScript often expose weaker types than they should. For
instance, the function \code{createElement} is polymorphic in its return type: it can return a \code{DivElement} as
well as an \code{InputElement}, among others, but the Dart, Fay, SharpKit and Kotlin APIs return the super-type of
all the possible values, namely the \code{Element} type. As a consequence, developers need to explicitly down-cast
the value they get, which is a tedious and error prone task.

Some other languages try to workaround this problem by using overloading instead of polymorphism. For instance, HaXe
provides functions \code{createDivElement}, \code{createInputElement}, which return a \code{DivElement} and an
\code{InputElement}, respectively. Besides requiring a higher effort to implement, this solution also reduces the
control level of users: by being statically resolved, the element type can not anymore be passed as a parameter.

It turns out that most of the existing statically typed languages compiling to JavaScript either loose control or
loose type safety when they expose Web browser’s APIs. How to give developers the same level of control as if they
were using the native Web APIs, but in a statically typed and convenient way?

In this paper we present several ways to integrate Web browser’s APIs as statically typed APIs that are safe and
give developers the same control level as if they were using the native APIs. We can achieve this by using advanced
features of type systems like dependent types and functional dependencies.

\section{Background Material}

Why is it difficult to type Web browser’s APIs?

\subsection{DOM Creation}

\begin{lstlisting}[language=JavaScript]
var div = document.createElement('div');
var input = document.createElement('input');
\end{lstlisting}

\code{div} has type \code{DivElement} while \code{input} has type \code{InputElement}. These types have different
methods. As stated in the introduction, most statically typed language have an API that returns an \code{Element},
which is the least upper bound of the possible types returned by \code{createElement}, forcing users to explicitly
down-cast the result to the expected type, thus loosing type safety.

Alternatively, some languages use overloading.

Consider the following function creating an element and giving it a class name:

\begin{lstlisting}[language=JavaScript]
var create = function (elementType, className) {
  val el = document.createElement(elementType);
  el.className = className;
  return el
};
create('input', 'field').focus();
create('img', 'figure').src = 'http://google.com/logo.png';
\end{lstlisting}

How to type check the above code? The problem is that the return type of the \code{create} function depends on its
first parameter.

Note that a possible solution in Java could be the following:

\begin{lstlisting}[language=Java]
class ElementName<E> {}
ElementName<InputElement> Input = new ElementName<InputElement>();
ElementName<ImageElement> Img = new ElementName<ImageElement>();

<E> E create(ElementName<E> type, String className) {
  ...
}

create(Input, "field").focus();
create(Img, "figure").setSrc("http://google.com/logo.png");
\end{lstlisting}

However, the type parameter \code{E} has to be filled at usage site. If the compiler can infer it, that’s fine.

\subsection{DOM Events}

A similar issue applies to the DOM events API.

\begin{lstlisting}[language=Java]
class EventName<E> {}
EventName<ClickEvent> Click = new EventName<ClickEvent>();

<E> Stream<E> on(EventName<E> type, Callback<E>) {
  ...
}

Callback<ClickEvent> incrementCounter = new Callback<ClickEvent>() {
  @Override
  void execute(ClickEvent event) {
    ...
  }
}

on(Click, incrementCounter);
\end{lstlisting}

Again, the type parameter \code{E} has to be filled at usage site.

\subsection{Selectors}

The problem with selectors is slightly different.

\begin{lstlisting}[language=JavaScript]
var input = document.querySelector('input');
var el = document.querySelector('#content');
\end{lstlisting}

\code{input} can not have another type than \code{InputElement}, but \code{el} can be an element of any type (depending
on which element has the id \code{content}). Return type can not \emph{always} be inferred from the parameters.
In this case, down-casting is impossible to avoid, but how to make it optional, and how to make it less risky?

We want to be able to write the following code:

\begin{lstlisting}
val input = document.find(Input) // input has type InputElement
val el = document.querySelector("#content") // el has type Element
val img = document.querySelector[Img](".figure") // note that the type Img is explicitly applied in this case but was automatically inferred in previous cases
val error = document.querySelector[Int]("div") // Does not compile: querySelector can not return a number
\end{lstlisting}

\section{Lightweight Modular Staging}

\section{Contribution}

\subsection{Events}

Path-dependent types to abstract over an event name and its data type.

\subsection{Selectors}

\begin{itemize}
 \item Less type annotations on DOM queries, less chance to write nonsense casts
 \item Inference-driving macros help inferring more specific types
\end{itemize}

\subsection{DOM}

?

\section{Evaluation}

\subsection{Events}

Other languages either provide loose information about the data type of the listened event (Dart) or give no way to
abstract over an event (GWT, Kotlin).

\section{Conclusion and Perspectives}

\begin{thebibliography}{4}

\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
Subsequences. J. Mol. Biol. 147, 195--197 (1981)

\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
Composing a Complex Biological Workflow through Web Services. In: Nagel,
W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
pp. 1148--1158. Springer, Heidelberg (2006)

\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
Infrastructure. Morgan Kaufmann, San Francisco (1999)

\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
Information Services for Distributed Resource Sharing. In: 10th IEEE
International Symposium on High Performance Distributed Computing, pp.
181--184. IEEE Press, New York (2001)

\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
Grid: an Open Grid Services Architecture for Distributed Systems
Integration. Technical report, Global Grid Forum (2002)

\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

\end{document}